---
title: "my_very_bad_intro"
author: "Eloise Whittle"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  word_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Heading 1 really big 

some text here 

i just love to write and talk (blah blah blah ... )


## Background 

some background text 

### Figure 1 

All good 

# line text formatting

**in bold** 
*italic* or _italic_

^superscript^

~subscript~

'inline' 

~~strikethrough~~

# line breaks 

this is my first time trying R markdown  
i have put two spaces, is this now on a new line?

now i am trying to make a new line using \
is this now on a new line?


# Links 

My paper has a Methods section. Please see here [link](#section-methods)



# Data amd methods{#section-methods}

## Link to a webpage 


[text of link](https://training.cam.ac.uk/bioinformatics/event-timetable)


## Footnotes 

[^1]: This is the first footnote  

[^2]: This is the second footnote


# Lists 

1. Item 1 
2. Item 2
3. Item 3
    a. item 3a 
        i. item 1
        ii. item ii
    b. item b 


## unordered 

* item 1 
* item 2 
    * item 3 
        * Item 4 
    * item 5 

- item 1 
    - item 1 
    
## Task lists 

- [ ] an unchecked task 
- [x] checked task 



# Inserting Images 

![Figure caption](path to image)




# Tables 

| Column 1    | Column 2    |
| ----------- | ----------- |
| Item 1,1    | Item 1,2    |
| Item 2,1    | Item 2,2    |

can create tables at [link of table generator](https://www.tablesgenerator.com/markdown_tables)  


| gene  | expression |
|-------|------------|
| BCL2  | 2.2        |
| MEF2b | 4.55       |
| TP53  | 8.90       |
| MYC   | 5.55       |



## Blocks 

> Example of a block 

if you want to insert code blocks 

```
print("Hello world")
x <- 1+2
print(x)
```

# Adding a table of contence 

output:
  html_document:
  toc: true 
  toc_float: true
 
at the beginning of the document under output.  





# finally some code 


```{r}
# heading 2 - this is not a heading it is a comment as we are not in the world of r. 
 cat("Hello World") 

plot(pressure)

``` 


```{r}
download.file(url="https://raw.githubusercontent.com/cambiotraining/reproducibility-training/master/data/gapminder_data.csv", destfile="data/gapminder_data.csv")
```

```{r, include = FALSE}

cat("hello World")

plot(pressure)
```
## Chunk lable 

```{r data=munging, each=false}

#load tidyverse 

library(tidyverse) #used for data manipulation
install.packages("rmarkdown")
library(rmarkdown) # used for paged_table function
install.packages("kableExtra")
library(kableExtra) # used for table
install.packages("ggpubr")
library("ggpubr") # used for ggarrange function

# read file into R 
library(readr)
pop_data <- read_csv("Data/gapminder_data.csv")

``` 

# plotting data and data munging

```{r}
pop_data

plot(pop_data$lifeExp, pop_data$gdpPercap)

```

## Filtering data

```{r}
# select only countries in Europe  
pop_data$continent == "Europe"

pop_data[1,2]

selected_frame <- pop_data[pop_data$continent == "Europe" & pop_data$year == 2007,]

euro_data_tbl <- pop_data %>%
  filter(continent == "Europ" & year == 2007) %>% 
  select(-continent, - year) %>%
  arrange(desc(lifeExp)) %>%
  rename(Country = country, "Populaation Size" = pop, 
         "Life Expectancy" = lifeExp, "GDP" = gdpPercap)

```


## Dynamic tables in R 

```{r}
head(selected_frame)

install.packages("kableExtra")
library(kableExtra)

selected_frame %>%
  kbl(caption = "Pop Data Selected") %>%
  kable_styling(bootstrap_options = c("striped", "hover", full_width = F, font = "Arial"))

```
## Challange 


```{r}
euro_data_fig <- pop_data %>%
  filter(continent == "Europe") %>%
  mutate(year = as.factor(year))

euro_plot <- ggplot(euro_data_fig, aes(x = year, y = lifeExp)) + geom_violin()
euro_plot

```


## arranging plots next to each other

```{r}
uk_data_fig <- pop_data %>%
  filter(country == "United Kingdom")

uk_plot <- ggplot(uk_data_fig, aes(y = lifeExp, x = year)) + geom_point()
uk_plot

ggarrange(euro_plot , uk_plot,labels = c("A", "B"))


```
